---
title: "WFAEx1CTA2024-02-06"
output: html_document
date: "2024-02-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages}
library(tidyverse)
library(tidytext) # note: not in tidyverse
library(ggthemes)
library(readr) #new way to import data
library(babynames) #for gender predictions
library("dplyr") 
library(magrittr)
```

##R Data

```{r dataimport}
edbfdata <- read_csv("https://raw.githubusercontent.com/cjbarrie/RDL-Ed/main/02-text-as-data/data/edbookfestall.csv")

```
First we check what we need, looking at the column names, and selecting relevant columns. 

````{r takedata}

colnames(edbfdata)
event_desc<-edbfdata %>%
  select(description, year)
colnames(event_desc)

````

We then find the number of events per year and plot it in ggplot: Note the use of "summarise" to create a new summary variable.

```{r plot}

events_year<-event_desc%>%
  mutate(obs=1) %>%
  group_by(year)%>%
  summarise(sum_events=sum(obs))
  
ggplot(events_year)+
  geom_line(aes(year,sum_events))+
    scale_y_continuous(expand = c(0, 0), limits = c(0, NA))

````

We are now going to tidy the text to get rid of html/php encodings as wella s punctuation, and stopwords
```{r tidytext}

tidy_des <- event_desc %>% 
  mutate(desc = tolower(description)) %>%
  unnest_tokens(word, desc) %>%
  filter(str_detect(word, "[a-z]"))%>%
  filter(!word %in% stop_words$word)

tidy_words1<-tidy_des%>%
  count(word, sort = TRUE)%>%
  head()



```

Using "head" we can see we need to take out rsquo, (which is an html encoding (for apostrophes)). however, there remain some encodings - we must deal with those as well.

```{r apostrophes}

tidy_des %>%
 filter(!word %in% c("rsquo"))%>%
head()

remove_reg <- c("&amp;","&lt;","&gt;","<p>", "</p>","&rsquo", "&lsquo;",  "&#39;", "<strong>", "</strong>", "rsquo", "em", "ndash", "nbsp", "lsquo", "strong")
                  

tidy_des %>%
 filter(!word %in% remove_reg)%>%
head()

tidy_des_words <- tidy_des %>%
  filter(!word %in% remove_reg)%>%
  count(word, sort = TRUE) %>%
  head()


````

Ok, now we have our words:

```{r counts}

edbf_term_counts <- tidy_des %>% 
  group_by(year) %>%
  count(word, sort = TRUE)

head(edbf_term_counts)

````
```{r changelistterm}
edbf_term_counts$womword <- as.integer(grepl("women|feminist|feminism|gender|harassment|sexism|sexist", x = edbf_term_counts$word))

head(edbf_term_counts)

```

````{r yearcounts}
edbf_counts <- edbf_term_counts %>%
  group_by(year) %>%
  mutate(year_total = sum(n)) %>%
  filter(womword==1) %>%
  summarise(sum_wom = sum(n),
            year_total= min(year_total))

head(edbf_counts)

```



```{r timetrends}
ggplot(edbf_counts, aes(year, sum_wom / year_total, group=1)) +
  geom_line() +
  geom_vline(xintercept = 2017, col="red")+
  geom_text(aes(x=2017.1, label="#metoo year", y=.0015),
            colour="black",angle=90, text=element_text(size=8))+
  xlab("Year") +
  ylab("% gender-related words") +
  scale_y_continuous(labels = scales::percent_format(),
                     expand = c(0, 0), limits = c(0, NA)) +
  theme_tufte(base_family = "Helvetica") 
````



